Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake

### 2014-10-14 paper.tex
Quick local.

\section{Implications for the Longmen Shan earthquake cycle}
The fundamental effect of allowing a 20 km detachment to accumulated and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip-deficit rate (\ref{fig:big_stack}a). Locking on the deep detachment obscures the signal of deformation on the range-front faults. The interseismic deformation is much more broadly distributed across more than 300 kilometers. According to this model, interseismically, the eastern Tibetan Plateau moves cohesively with the Sichuan Basin, whereas coseismic slip is concentrated at the boundary between these two blocks.

@@

While we have explained the observed the broad GPS interseismic velocity gradient as a result of a locked fault geometry including a deep detachment, we have not addressed when accumulated slip-deficit on the deep detachment is released. Geodetically constrained estimates of co- and post-seismic slip shows up to 5 m of slip on a deep detachment \citep{Qi2011}. Our data analysis depends on the detachment being locked during the period of data collection from 1998 to 2004 \citep{gan07}. In addition to coseismic and short-term postseismic activity on the detachment, it is possible that intermittent aseismic slip could be an important mode of strain release. This is difficult to assess given that intersimsic GPS observations cover, at best, 6\% of a complete Longmen Shan earthquake cycle. Seismicity is approximately limited to the upper 20 km of crust in the range-front \citep{Li2010a}. Using steady-state critical taper wedge theory \citep{dahlen90}, with a internal coefficient of friction of 0.7, a cohesion of 5 MPa, and a hydrostatic fluid pressure, we find a basal friction coefficient of 0.08. This is comparable to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods. On the other hand, the surface area and shortening rate of the detachment could sustain very large earthquakes if slip is released coseismically. Specifically, if 100 km down-dip and 300 km along-strike were to slip 4 meters, similar to average slip during the Wenchuan earthquake, a $M_{\textrm{w}}$ 8.4 rupture would be produced.

@@

The 6-9 mm/yr slip-deficit rates estimated here can be used to provide new constraints on earthquake recurrence intervals at the Longmen Shan (Figure \ref{fig:hazard}). Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years, with much shorter recurrence intervals for smaller events. This contrasts with the average recurrence of ~2000 years inferred from paleoseismic field investigations for the Beichuan fault \citep{Ran2010}. The 6-9 mm/yr slip-deficit rate estimated here would imply $M_{\textrm{w}} = $9.0 ruptures for that recurrence interval, which would seemingly require an along strike rupture equal to or greater than the ${\sim}300$ km along strike length of the Longmen Shan. However, our slip-deficit estimate could include the effects of other imbricated structures. As a result our estimate is a system-wide estimate while paleoseismic estimates have focused on the analysis of individual thrust faults.

>>>>

\section{Implications for the Longmen Shan earthquake cycle}
The fundamental effect of allowing a 20 km detachment to accumulate and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip-deficit rate (\ref{fig:big_stack}a). Locking on the deep detachment obscures the signal of deformation on the range-front faults. The interseismic deformation is much more broadly distributed across more than 300 kilometers. According to this model, interseismically, the eastern Tibetan Plateau moves cohesively with the Sichuan Basin, whereas coseismic slip is concentrated at the boundary between these two blocks.

@@

While we have explained the observed the broad GPS interseismic velocity gradient as a result of a locked fault geometry including a deep detachment, we have not addressed when accumulated slip-deficit on the deep detachment is released. Geodetically constrained estimates of co- and post-seismic slip shows up to 5 m of slip on a deep detachment \citep{Qi2011}. Our data analysis depends on the detachment being locked during the period of data collection from 1998 to 2004 \citep{gan07}. In addition to coseismic and short-term postseismic activity on the detachment, it is possible that intermittent aseismic slip could be an important mode of strain release. This is difficult to assess given that intersimsic GPS observations cover, at best, 6\% of a complete Longmen Shan earthquake cycle. Seismicity is approximately limited to the upper 20 km of crust in the range-front \citep{Li2010a}. Using steady-state critical taper wedge theory \citep{dahlen90}, with an internal coefficient of friction of 0.7, a cohesion of 5 MPa, and a hydrostatic fluid pressure, we find a basal friction coefficient of 0.08. This is comparable to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods. On the other hand, the surface area and shortening rate of the detachment could sustain very large earthquakes if slip is released coseismically. Specifically, if 100 km down-dip and 300 km along-strike were to slip 4 meters, similar to average slip during the Wenchuan earthquake, a $M_{\textrm{w}}$ 8.4 rupture would be produced.

@@

The 6-9 mm/yr slip-deficit rates estimated here can be used to provide new constraints on earthquake recurrence intervals at the Longmen Shan (Figure \ref{fig:hazard}). Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years, with much shorter recurrence intervals for smaller events. This contrasts with the average recurrence of ~2000 years inferred from paleoseismic field investigations for the Beichuan fault \citep{Ran2010}. The 6-9 mm/yr slip-deficit rate estimated here would imply $M_{\textrm{w}} = $9.0 ruptures for that recurrence interval, which would seemingly require an along strike rupture equal to or greater than the ${\sim}300$ km along strike length of the Longmen Shan. However, our slip-deficit estimate could include the effects of other imbricated structures. As a result our estimate is system-wide while paleoseismic estimates have focused on the analysis of individual thrust faults.
